VM-Centric Snapshot Deduplication\\ for Cloud Data Backup

### 2015-03-14 snapshot_system/abs.tex
re

\begin{abstract}
A cloud environment that  hosts a large number of  virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots and
signature-based deduplication of data blocks is necessary to eliminate excessive 
redundant blocks.  Collocating a cluster-based duplicate service with other cloud services
reduces network traffic;
% when their signatures are identical. 
however it is resource expensive and less fault-resilient to perform a global deduplication
and let a data block share by many virtual machines. 
This paper proposes a VM-centric collocated backup service that  
localizes deduplication as much as possible within each virtual machine using similarity-guided
search and associates  underlying file blocks with one VM for most cases.
It restricts cross-VM deduplication under common blocks with extra replication support.
Our analysis shows that  this VM centric scheme 
can  provide  better fault tolerance while using a small amount of computing and storage resources. 
This  paper  describes a comparative evaluation of this scheme in  accomplishing a competitive deduplication 
efficiency while sustaining a good backup throughput.

>>>>

\begin{abstract}
A cloud environment that  hosts a large number of  virtual machines (VMs) has
a high storage demand for frequent backup of system image snapshots and
signature-based deduplication of data blocks is necessary to eliminate excessive 
redundant blocks.  Collocating a cluster-based duplicate service with other cloud services
reduces network traffic;
% when their signatures are identical. 
however it is resource expensive and less fault-resilient to perform a global deduplication
and let a data block share by many virtual machines. 
This paper proposes a VM-centric collocated backup service that  
localizes deduplication as much as possible within each virtual machine using similarity-guided
search and associates  underlying file blocks with one VM for most cases.
It restricts cross-VM deduplication under common blocks with extra replication support.
This  paper  describes an analytical and experimental evaluation of this scheme 
in providing  better fault tolerance and competitive deduplication efficiency
while using a small amount of computing and storage resources. 
%for  accomplishing a competitive deduplication 
%efficiency while sustaining a good backup throughput.

### 2015-03-14 snapshot_system/mass_abs.tex
re

\begin{abstract}
Data deduplication is important  for snapshot backup of virtual machines (VMs)
because of excessive redundant content.  Fingerprint search for source-side duplicate 
detection is resource intensive when the backup service for VMs  is co-located with  other 
cloud services.  This paper presents the design and implementation of
a fast VM-centric backup service with  a tradeoff 
for a competitive deduplication efficiency while using small computing resources, 
suitable for running on a converged cloud architecture that cohosts  many other services.
The  developed optimization includes similarity-guided local search 
for  the improved overall deduplication effectiveness 
and  VM-centric file system block management for the increased VM snapshot availability.  
%deduplication under popular chunks with extra replication support.
%and use an approximate method for fast and simplified snapshot deletion.
%by separating popular chunks and associating  file system blocks with one VM for non-popular chunks.
This  paper  describes an evaluation of this VM-centric scheme to assess  its deduplication 
efficiency, resource usage,  and fault tolerance.

>>>>

\begin{abstract}
Data deduplication is important  for snapshot backup of virtual machines (VMs)
because of excessive redundant content.  Fingerprint search for source-side duplicate 
detection is resource intensive when the backup service for VMs  is co-located with  other 
cloud services.  This paper presents the design and analysis of
a fast VM-centric backup service with  a tradeoff 
for a competitive deduplication efficiency while using small computing resources, 
suitable for running on a converged cloud architecture that cohosts  many other services.
The  design consideration includes 
VM-centric file system block management for the increased VM snapshot availability.
%and  similarity-guided local search for  the improved overall deduplication effectiveness. 
%deduplication under popular chunks with extra replication support.
%and use an approximate method for fast and simplified snapshot deletion.
%by separating popular chunks and associating  file system blocks with one VM for non-popular chunks.
This  paper  describes an evaluation of this VM-centric scheme 
to assess  its deduplication efficiency, resource usage,  and fault tolerance.

### 2015-03-14 snapshot_system/mass_conc.tex
re

\section{Conclusions}
\label{sect:conclusion}
%In this paper we propose a collocated backup service built on
%the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The main contribution of this paper is the development and comparative study of a low-profile and VM-centric deduplication scheme that
uses  a small amount of system resource in both snapshot deduplication and deletion
while delivering competitive deduplication efficiency.
The VM-centric design also allows an optimization for  better fault isolation and tolerance.
Our evaluation has the following results.
\begin{itemize}
\item {\em Competitive deduplication efficiency with a tradeoff}. 
The proposed VC scheme 
can accomplish over  96\% of what complete global
deduplication can do. In comparison, SRB~\cite{Dong2011,extreme_binning09}
can accomplish 97.86\% while 
the sampled index~\cite{Guo2011} can deliver 97\% for a different test dataset. 
\item {\em Lower resource usage in deduplication}. 
The VC scheme achieves a 85K:1 ratio between raw snapshot data and memory usage
and is much more memory-efficient than 
SRB  with 4K:1 ratio and sampled index with 20K:1.
VC with 100 machines takes 1.31 hours to complete the backup of 2,500 VMs
using 19\% of one core and 16.7\% of IO bandwidth per machine. 
Processing time of VC is 23.9\% faster  than SRB in our tested cases
and the per-machine throughput of VC is reasonable based on the result 
from~\cite{Guo2011}. 
Noted that both VC and SRB are designed for a distributed cluster architecture while
the sampled index method is for a single-machine architecture.
\item {\em Lower cost in snapshot deletion}. 
The deletion in VC with summary vectors
takes less than a minute for snapshot deletion using 15MB memory per machine
and all machines can run this operation in parallel without data dependence. 
The grouped mark-sweep method can take 43 hours using
1.2 to 3GB memory per machine in our tested cases.
The leak repair in VC takes 0.82 hours using 50MB on average while processing
some VMs may use upto 1.96GB. Such a repair is conducted infrequently (e.g. every 19.6 days) 
\item {\em Higher availability}. 
%The availability of snapshots increases substantially when
%adding more replication for popular cross-VM file blocks and packaging chunks from the same 
%VM in one file system block. 
The snapshot availability of VC is 99.66\% with 3 failed machines in a 100-machine cluster
while it is 69.54\% for  SRB or a VM-oblivious approach.
The analysis shows that the replication degree
for the popular data set between 6 and 9 is good enough when the replication degree
for other data blocks is 3, and adds only a small cost to storage.
\end{itemize}
The offline PDS recomputation does require some modest I/O and memory resource and 
since the recomputing frequency is relatively low (e.g. on a monthly basis), we expect such resource
consumption is acceptable in practice.

>>>>

\section{Discussions and Conclusions}
\label{sect:conclusion}
%In this paper we propose a collocated backup service built on
%the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The main contribution of this paper is the development and analysis of a low-profile and VM-centric deduplication scheme that
uses  a small amount of system resource in both snapshot deduplication and deletion
while delivering competitive deduplication efficiency.
The VM-centric design also allows an optimization for  better fault isolation and tolerance.
Our evaluation has the following results.
\begin{itemize}
\item {\em Competitive deduplication efficiency with a tradeoff}. 
The proposed VC scheme 
can accomplish over  96\% of what complete global
deduplication can do. In comparison, SRB~\cite{Dong2011,extreme_binning09}
can accomplish 97.86\% while 
the sampled index~\cite{Guo2011} can deliver 97\% for a different test dataset. 
Local similarity search removes 12.05\% of the original data after using a segment-based
 dirty bit method and among them, 3\% is contributed by the increased similarity search compared to
a simple dirty offset method in ~\cite{WeiZhangIEEE}.
\item {\em Lower resource usage in deduplication}. 
The VC scheme achieves a 85K:1 ratio between raw snapshot data and memory usage
and is much more memory-efficient than 
SRB  with 4K:1 ratio and sampled index with 20K:1.
VC with 100 machines takes 1.31 hours to complete the backup of 2,500 VMs
using 19\% of one core and 16.7\% of IO bandwidth per machine. 
Processing time of VC is 23.9\% faster  than SRB in our tested cases
and the per-machine throughput of VC is reasonable based on the result 
from~\cite{Guo2011}. 
Noted that both VC and SRB are designed for a distributed cluster architecture while
the sampled index method is for a single-machine architecture.
\item {\em Higher availability}. 
%The availability of snapshots increases substantially when
%adding more replication for popular cross-VM file blocks and packaging chunks from the same 
%VM in one file system block. 
The snapshot availability of VC is 99.66\% with 3 failed machines in a 100-machine cluster
while it is 69.54\% for  SRB or a VM-oblivious approach.
The analysis shows that the replication degree
for the popular data set between 6 and 9 is good enough when the replication degree
for other data blocks is 3, and adds only a small cost to storage.
\end{itemize}
The offline PDS recomputation does require some modest I/O and memory resource and 
since the recomputing frequency is relatively low (e.g. on a monthly basis), we expect such resource
consumption is acceptable in practice.

%We have also developed a VM-centric strategy for snapshot deletion and the VM-centric design
%allows the deletion be localized for each VM, which greatly simplifies the reference counting
%process. That is beyond  the scope of this paper.

\comments{
\item {\em Lower cost in snapshot deletion}. 
The deletion in VC with summary vectors
takes less than a minute for snapshot deletion using 15MB memory per machine
and all machines can run this operation in parallel without data dependence. 
The grouped mark-sweep method can take 43 hours using
1.2 to 3GB memory per machine in our tested cases.
The leak repair in VC takes 0.82 hours using 50MB on average while processing
some VMs may use upto 1.96GB. Such a repair is conducted infrequently (e.g. every 19.6 days) 
}

### 2015-03-14 snapshot_system/mass_dedup.tex
re

%when restricting fingerprint comparison within a VM uses
%only a small amount of resource. 
%ehat the local deduplication has removed most of the duplicates.
%There are fewer deduplication opportunities across VMs while the memory and network
%consumption for global comparison is more expensive.
%Thus our approximation is that the global cross-VM fingerprint comparison only searches for the top $k$
%most popular items. 
PDS  can be computed in an offline manner periodically, e.g., on a monthly basis.
%Once the popularity of all data chunks is collected, the system only maintains the top $k$
%most popular chunk fingerprints (called \textbf{PDS index}) in a distributed shared memory.  
%These top chunks are shared among multiple VMs and  since $\sigma$ is relatively small, 
%we can afford to provide extra replicas for these popular chunk data to enhance the fault resilience.
%Compared to a preliminary solution which  uses data popularity~\cite{WeiZhangIEEE}, 
%this paper provides  a more comprehensive scheme with  improved deduplication efficiency and fault tolerance, and
%analytic design guidance. 
% Figure~\ref{fig:Datazipf} illustrates the distribution of chunk popularity based on a data trace from Alibaba with over 2500 VMs.
% The distribution is zipf like and popular chunks are dominating the large portion of data chunks.
% We denote $\sigma$ as the percentage of unique data chunks belonging to PDS and from the evaluation we find that
% $\sigma$ with a range of 1 to 4\% can deliver a fairly competitive deduplication efficiency.
Compared to ~\cite{bottleneck08}, the fingerprint index  size is reduced by $100-\sigma$ percentage
(e.g. 98\% with $\sigma=2\%$)
and the searchable fingerprint space  becomes very small under the popularity constraint.  
The fingerprint-guided distributed mapping in ~\cite{extreme_binning09,Dong2011} narrows
the search scope of each data chunk, but it does not reduce th total amount of searchable fingerprints
used for the entire deduplication. 
%In comparison we only focus on top popular chunks among VMs and this
%significantly reduces the total amount of  searchable fingerprints.

@@

Since the previous work typically uses a non-uniform chunk size 
with an average of 4KB or 8KB for the best deduplication 
effectiveness~\cite{Guo2011,extreme_binning09,bottleneck08,Dong2011},
we conduct additional local similarity guided deduplication 
by comparing chunk fingerprints of a dirty segment in a snapshot 
with those in  {\em potentially similar} segments from its parent snapshot. 
We consider a parent  segment is  potentially similar to the current dirty segment if 1) the parent segment
is at the same offset as the current segment.
2) the signature of these segments is the same.
The signature of a segment is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). 
%Note that this definition of content similarity is an approximation~\cite{resemblance97}.  
This signature definition is motivated by the previous work for similarity-based routing in parallel
deduplication~\cite{extreme_binning09,Dong2011} and near duplicate detection~\cite{shingling97}.

>>>>

%when restricting fingerprint comparison within a VM uses
%only a small amount of resource. 
%ehat the local deduplication has removed most of the duplicates.
%There are fewer deduplication opportunities across VMs while the memory and network
%consumption for global comparison is more expensive.
%Thus our approximation is that the global cross-VM fingerprint comparison only searches for the top $k$
%most popular items. 
PDS  can be computed in an offline manner periodically, e.g., on a monthly basis.
%Once the popularity of all data chunks is collected, the system only maintains the top $k$
%most popular chunk fingerprints (called \textbf{PDS index}) in a distributed shared memory.  
%These top chunks are shared among multiple VMs and  since $\sigma$ is relatively small, 
%we can afford to provide extra replicas for these popular chunk data to enhance the fault resilience.
%Compared to a preliminary solution which  uses data popularity~\cite{WeiZhangIEEE}, 
%this paper provides  a more comprehensive scheme with  improved deduplication efficiency and fault tolerance, and
%analytic design guidance. 
% Figure~\ref{fig:Datazipf} illustrates the distribution of chunk popularity based on a data trace from Alibaba with over 2500 VMs.
% The distribution is zipf like and popular chunks are dominating the large portion of data chunks.
% We denote $\sigma$ as the percentage of unique data chunks belonging to PDS and from the evaluation we find that
% $\sigma$ with a range of 1 to 4\% can deliver a fairly competitive deduplication efficiency.
Compared to ~\cite{bottleneck08}, the fingerprint index  size is reduced by $100-\sigma$ percentage
(e.g. 98\% with $\sigma=2\%$)
and the searchable fingerprint space  becomes very small under the popularity constraint.  
The fingerprint-guided distributed mapping in ~\cite{extreme_binning09,Dong2011} narrows
the search scope of each data chunk, but it does not reduce th total amount of searchable fingerprints
used for the entire deduplication. 
%In comparison we only focus on top popular chunks among VMs and this
%significantly reduces the total amount of  searchable fingerprints. 
Appendix~\ref{sect:dedupanalysis} provides an analysis of the $k$ value selection and
PDS effectiveness.

%under this VC scheme and compares it with VO.

@@

Since the previous work typically uses a non-uniform chunk size 
with an average of 4KB or 8KB for the best deduplication 
effectiveness~\cite{Guo2011,extreme_binning09,bottleneck08,Dong2011},
we conduct additional local similarity guided deduplication 
by comparing chunk fingerprints of a dirty segment in a snapshot 
with those in  {\em potentially similar} segments from its parent snapshot. 
We consider a parent  segment is  potentially similar to the current dirty segment if 1) the parent segment
is at the same offset as the current segment.
2) the signature of these segments is the same.
The signature of a segment is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). 
%Note that this definition of content similarity is an approximation~\cite{resemblance97}.  
This above approach is motivated by the previous work for similarity-based 
deduplication~\cite{shingling97,extreme_binning09,Dong2011,Aronovich2009} while our focus is local search.

### 2015-03-14 snapshot_system/mass_intro.tex
re

%Without deduplication,
Source-side deduplication (e.g. ~\cite{Symantec2009})
 eliminates duplicates before backup data is transmitted
to a secondary storage, which saves network bandwidth significantly; however its resource
usage can impact other co-located cloud services. 
It is memory-intensive to compare a large number of fingerprints and identify
duplicates, even with optimization or approximation techniques developed~\cite{Guo2011,Dong2011,extreme_binning09}. 
\comments{
When deleting unused snapshots, a grouped mark-and-sweep approach~\cite{Guo2011}  
is effective while it still carries a significant cost, especially in a distributed
setting. 
} Another side effect of deduplication is the possible loss of failure
resilience~\cite{Reliability06}.  
%Separate files share the same physical copies of blocks that are
%logically duplicated among them.  
If a shared block is lost, all files that
share that block are affected.  
A cloud may offload backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out raw and undeduplicated backup data wastes a huge amount
of network bandwidth that would otherwise be available to user VMs. 
\comments{
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).
}

@@

The main contribution of this paper is
1) a similarity-guided local deduplication for
each VM to make the overall deduplication efficiency more competitive 
to a transitional deduplication scheme;
2) a VM-centric file block management to improve snapshot availability by
separating popular chunk replication and
packaging data chunks from the same VM into a file system block as much as possible; 
3) a design and implementation of a VM-centric  backup service to integrate and validate
the proposed techniques while imposing  little resource usage and  minimizing the impact on other 
collocated cloud services.We provide an analysis on the deduplication efficiency by using popular items
and also on fault isolation.
Our  experimental study compares  the VM-centric approach with the previous 
VM-oblivious approaches in terms of deduplication efficiency, resource usage,  and snapshot availability.
%Thus we improve  to avoid such VM-oblivious packaging.
%our approach uses a deduplication data chunk size that is smaller than a
%single file block and

>>>>

%Without deduplication,
Source-side deduplication (e.g. ~\cite{Symantec2009})
 eliminates duplicates before backup data is transmitted
to a secondary storage, which saves network bandwidth significantly; however its resource
usage can impact other co-located cloud services. 
It is memory-intensive to compare a large number of fingerprints and identify
duplicates, even with optimization or approximation techniques developed~\cite{Guo2011,Dong2011,extreme_binning09,FuFast2015}. 
\comments{
When deleting unused snapshots, a grouped mark-and-sweep approach~\cite{Guo2011}  
is effective while it still carries a significant cost, especially in a distributed
setting. 
} Another side effect of deduplication is the possible loss of failure
resilience~\cite{Reliability06}.  
%Separate files share the same physical copies of blocks that are
%logically duplicated among them.  
If a shared block is lost, all files that
share that block are affected.  
A cloud may offload backup workload from production 
server hosts to dedicated backup proxy servers (e.g. EMC Data Domain) or
backup services (e.g. Amazon S3). This approach simplifies the cluster 
architecture and 
avoids potential performance degradation to production applications when 
backups are in progress.
However, sending out raw and undeduplicated backup data wastes a huge amount
of network bandwidth that would otherwise be available to user VMs. 
\comments{
Even when the storage system implements deduplication internally
such dedicated backup storage is either difficult to scale out
(which can comprise the overall scalability of the converged architecture) 
or it must forego global deduplication in order to gain scalability 
(which increases backup cost).
}

@@

The main contribution of this paper is
1) a design and implementation of a backup service 
to integrate the VM-centric techniques while minimizing the resource impact on other 
collocated cloud services;
2) a VM-centric file block management to improve snapshot availability by
separating popular chunk replication and
packaging data chunks from the same VM into a file system block as much as possible; 
3) an analysis on the deduplication efficiency by using popular items
and also on the impact of above fault isolation design on snapshot availability.
The integrated optimization includes  similarity-guided local deduplication for
each VM to make the overall deduplication efficiency more competitive 
to a traditional deduplication scheme.
Our  experimental study compares  the VM-centric approach with the previous 
VM-oblivious approaches in terms of deduplication efficiency, resource usage,  and snapshot availability.
%Thus we improve  to avoid such VM-oblivious packaging.
%our approach uses a deduplication data chunk size that is smaller than a
%single file block and
